Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-02 overleaf/paper_SIMD.tex
Cut more.
Finally

@@

We presented a formalization of the MPC Source intermediate language followed by a specific back-end optimization at the level of MPC Source: novel SIMD-vectorization. We demonstrated that vectorization has significant impact on performance. 
We are excited about the opportunities for future work --- integration with protocol mixing, divide-and-conquer reasoning and parallelization, as well as inter-procedural context-sensitive analysis at the level of MPC Source will improve MPC programmability and efficiency.

>>>>

\newcommand{\squeeze}{\vspace{-0.2cm}}

@@

We presented a formalization of the MPC Source intermediate language followed by a specific back-end optimization at the level of MPC Source: novel SIMD-vectorization. We demonstrated that vectorization has significant impact on performance. 
We are excited about many opportunities for future work --- integration with protocol mixing, divide-and-conquer reasoning and parallelization, as well as inter-procedural context-sensitive analysis for  MPC Source. %will improve MPC programmability and efficiency.

### 2022-05-02 overleaf/related.tex
Finally

\paragraph{MPC languages and compilers.}
Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
These two families are the state-of-the art and
are actively developed. Another recent development is Viaduct~\cite{Acay:2021}, a functional language and compiler that supports a range of secure computation
frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}. We believe that our work is unique as 
as a facade between the high-level language and lower-level circuit protocols.

@@

HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY circuits. It does source-to-source compilation with the key goal to decompose the 
program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
future optimizations as well. We formalize MPC Source and reasoning about transformations, which we conjecture is more tractable than reasoning over
the higher-level AST. On the other hand, HyCC does inter-procedural optimizations, while our analysis is intra-procedural. We will explore context-sensitive inter-procedural 
analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
Par4All (\url{https://github.com/Par4All/par4all}), however, does not appear to be included with the publicly available distribution of HyCC.}
to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
problem than what MPC presents and may forgo opportunities for optimization --- concretely, it is well-known that vectorization and polyhedral compilation 
do not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops and achieve significant speedup.

%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
\squeeze
\paragraph{Classical HPC compilers.}
Automatic vectorization is a longstanding problem in high-performance computing and
there are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to
MPC Source vectorization.
We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes
both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the
result of the condition at runtime.
The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
can lead to slowdown.
%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}

We believe that vectorization over linear MPC Source warrants a new look, while drawing from
results in HPC.
%Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
%we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
%Vectorization meshes in with those additional optimizations in non-trivial ways.
%Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
%and present novel challenges and opportunities for contribution. 
Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
about dependence, which we conjecture is more suitable for MPC Source. 
%Buscher~\cite{Buscher:2018} considers SIMD-vectorization
%at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
%He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
%essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
%In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
%and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.

>>>>

\paragraph{MPC languages and compilers.}
Languages and compilers for secure computation have seen significant advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
These two families are the state-of-the art and
are actively developed. Another recent development is Viaduct~\cite{Acay:2021}, a functional language and compiler that supports a range of secure computation
frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}. We believe that our work is unique as 
as it falls between the high-level language and lower-level circuit protocols.

@@

HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY circuits. It does source-to-source compilation with the key goal to decompose the 
program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
future optimizations as well. We formalize MPC Source and reasoning about transformations, which we conjecture is more tractable than reasoning over
the higher-level AST. On the other hand, HyCC does inter-procedural optimizations, while our analysis is intra-procedural. We will explore context-sensitive inter-procedural 
analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
Par4All (\url{https://github.com/Par4All/par4all}), however, does not appear to be included with the publicly available distribution of HyCC.}
to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
problem than what MPC presents and may forgo opportunities for optimization --- concretely, it is well-known that vectorization and polyhedral compilation 
do not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops. %and achieve significant speedup.

%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
\squeeze
\paragraph{Classical HPC compilers.}
Automatic vectorization is a longstanding problem in high-performance computing and
there are thousands of works reflecting many years of research. We presented a vectorization
algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to
MPC Source vectorization.
We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes
both branches of a conditional applying masking to avoid execution of the branch-not-taken code as well as selection (similar to MUX). 
%to select the correct value based on the
%result of the condition at runtime.
%The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
%can lead to slowdown.
%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}

We believe that vectorization over linear MPC Source warrants a new look, while drawing from
results in HPC.
%Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
%we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
%Vectorization meshes in with those additional optimizations in non-trivial ways.
%Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
%and present novel challenges and opportunities for contribution. 
Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
vectorization over SSA form. %, which has similarities to MPC Source. 
We consider different array representation, notion of dependence, and reasoning
about dependence, which is more suitable for MPC Source. 
%Buscher~\cite{Buscher:2018} considers SIMD-vectorization
%at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
%He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
%essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
%In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
%and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.

### 2022-05-02 overleaf/results.tex
More cuts.
Finally

\subsection{Experiment Setup}\label{sec:experiment_setup}
We tested our framework with several benchmarks. For the multiparty computation (MPC), we restricted our evaluation to 2 party computation (2PC) setting because it requires fewer computing resources. We stress that there is no such inherent restriction in our framework. We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Network (LAN) and Wide Area Network (WAN). In the LAN setting, we use {\tt c6525-25g} machines for both parties. These machines are equipped with 16-core AMD 7302P 3.0GHz processors and 128GB of RAM. The connection between these machines had 10Gbps bandwidth and sub-millisecond latency. %This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks. 
In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second. The {\tt c220g1} machine is equipped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM. We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms. %At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.

We run all experiments 5 times and report average values of various metrics. Note that standard deviation --- shown as error bar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value. %,  therefore the accuracy of the results is not effected by the relatively fewer runs.

@@

We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynth~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols). 
We did not use Gcd and Modular Exponentiation from~\cite{Ishaq:2019} as they do not benefit from vectorization according to our model. 
We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
in our syntax at the moment and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
The extended version describes what the benchmarks do in detail~\cite{Anon_TR}.

In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments. The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
% however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).

@@

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
\caption{Biometric Matching Circuit Evaluation Time, x-axis lists database size}
\label{fig:graph_biometric_eval_time}
\end{figure}

@@

%We conclude the . 
Finally, we compared our automatically generated routine for Inner Product with the manually SIMD-ified MOTION-native routine in the distribution. 
We were surprised that we were an order of magnitude slower in Boolean GMW as our circuit ran a significantly larger 
number of communication rounds. Upon investigation, it turned out that the vectorized multiplications were essentially the same, 
however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
{\sf result += mult\_unsimdified[i];} while our loop ran auto generated {\sf result[i] = result\_prev[i] + mult\_unsimdified[i];} (recall that scalar expansion
is an artifact of our vectorization). We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!

MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to a
limitation of our current implementation. Recall that the Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above slightly more complex code. 
Our experience indicates the limitation of AST-based analysis; because AST analysis is difficult it does not appear to detect the associative accumulation.
%(unfortunately, we ran out of time and did not experiment further).
%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.

@@

The most closely related work, HyCC~\cite{CCS:BDKKS18}, is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
Again, we estimate we are about an order of magnitude slower, which is likely due to the same issue as Inner Product --- the computation of $\mathit{min}$ can be optimized.

>>>>

\subsection{Experiment Setup}\label{sec:experiment_setup}
We tested our framework with several benchmarks. For the multiparty computation (MPC), we restricted our evaluation to 2 party computation (2PC) setting because it requires fewer computing resources. We stress that there is no such inherent restriction in our framework. We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Network (LAN) and Wide Area Network (WAN). In the LAN setting, we use {\tt c6525-25g} machines for both parties. These machines are equipped with 16-core AMD 7302P 3.0GHz processors and 128GB of RAM. The connection between these machines had 10Gbps bandwidth and sub-millisecond latency. %This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks. 
In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second. The {\tt c220g1} machine is equipped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM. We measured the connection bandwidth to be 560Mbps and average round trip time (RTT) to be 38ms. %At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.

We run all experiments 5 times and report average values of various metrics. Note that standard deviation in all observations was at most 4.5\% of the mean value. %,  therefore the accuracy of the results is not effected by the relatively fewer runs.

@@

We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynth~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols). 
We did not use Gcd and Modular Exponentiation from~\cite{Ishaq:2019} as they do not benefit from vectorization according to our model. 
We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
\cite{Anon_TR} describes what the benchmarks do.

To compare we run {\em both} versions with the maximal value of $N$ for which the non-vectorized version succeeds (typically {\tt N=128}) and we run {\em vec} with typically {\tt N=4096}.
Again, due to space constraints \cite{Anon_TR} gives the exact values.
% In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
%In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments. The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
% however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).

@@

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
\caption{Biometric Matching Circuit Evaluation Time}
\label{fig:graph_biometric_eval_time}
\end{figure}

@@

%We conclude the . 
Finally, we compared our automatically generated routine for Inner Product with the manually SIMD-ified MOTION-native routine in the distribution. 
We were surprised that we were an order of magnitude slower in Boolean GMW as our circuit ran a significantly larger 
number of communication rounds. Upon investigation, it turned out that the vectorized multiplications were essentially the same, 
however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
{\sf result += mult\_unsimdified[i];} while our loop ran auto generated {\sf result[i] = result\_prev[i] + mult\_unsimdified[i];} (as scalar expansion
is an artifact of our vectorization). We rewrote the accumulation manually and that led to the comparable speedup!

MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to
limitations of our implementation. Recall that Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that static analysis 
will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above slightly more complex code. 
However, because AST analysis is difficult it does not appear to detect the accumulation.
%(unfortunately, we ran out of time and did not experiment further).
%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.

@@

HyCC~\cite{CCS:BDKKS18} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ as its focus is mixing. %; it does provide a lot of data on mixed protocol executions.
We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.

### 2022-05-02 overleaf/vectorization.tex
Finally

Phases 3 cleans up local arrays of references (this is an optional phase and our current implementation does not include it)
and Phase 4 explicitly turns operations into MOTION SIMD operations.

@@

\COMMENT { Phase 4: }
\STATE {add SIMD for simdified dimensions}

@@

\paragraph{MPC Source Syntax}

\figref{fig:mpc_syntax} states the syntax and linearization semantics of MPC Source. 
Although notation is heavy, the linearization simply produces schedules as discussed in~\secref{sec:model}; 
the iterative MPC Source gives rise to what we called sequential schedule where loops are unrolled 
and MPC Source with vectorized dimensions gives rise to what we called parallel schedule. 
For simplicity, we consider only scalars and read-only arrays, however, the treatment extends to write arrays as well.
$\x[i,J,k]$ denotes the value of scalar variable \x{} at loop nest $i,j,k$. 
%(We use 3 dimensions, $i,j,k$, however, the treatment generalizes to arbitrary dimensions.) 
Upper case $J$ denotes a vectorized dimension and lower case $i,k$ denote iterative dimensions. Our compiler imposes 
semantic restrictions over the syntax: (1) $\x$ is treated as a 3-dimensional array and (2) $\x[i,J,k]$ must be enclosed
into for-loops on non-vectorized dimensions $i$ and $k$:
\begin{pythonn}
for i in range(I):
   ...
   for k in range(K):
      ... x[i,J,k] ...
\end{pythonn}

@@

\paragraph{Partial Orders}

@@

\paragraph{Theorems} The two theorems arising from the Basic vectorization optimization are as follows:

>>>>

Phases 3 cleans up local arrays of references (this is an optional phase and our current implementation does not include it).
%and Phase 4 explicitly turns operations into MOTION SIMD operations.

@@

%\COMMENT { Phase 4: }
%\STATE {add SIMD for simdified dimensions}

@@

\squeeze
\paragraph{MPC Source Syntax}

\figref{fig:mpc_syntax} states the syntax and linearization semantics of MPC Source. 
Although notation is heavy, the linearization simply produces schedules as discussed in~\secref{sec:model}; 
the iterative MPC Source gives rise to what we called sequential schedule where loops are unrolled 
and MPC Source with vectorized dimensions gives rise to what we called parallel schedule. 
For simplicity, we consider only scalars and read-only arrays, however, the treatment extends to write arrays as well.
$\x[i,J,k]$ denotes the value of scalar variable \x{} at loop nest $i,j,k$. 
%(We use 3 dimensions, $i,j,k$, however, the treatment generalizes to arbitrary dimensions.) 
Upper case $J$ denotes a vectorized dimension and lower case $i,k$ denote iterative dimensions. Our compiler imposes 
semantic restrictions over the syntax: (1) $\x$ is treated as a 3-dimensional array and (2) $\x[i,J,k]$ must be enclosed
into for-loops on non-vectorized dimensions $i$ and $k$:
\begin{pythonn}
for i in range(I):
   for k in range(K):
      ... x[i,J,k] ...
\end{pythonn}

@@

\squeeze
\paragraph{Partial Orders}

@@

\squeeze 
\paragraph{Theorems} The two theorems arising from the Basic vectorization optimization are as follows:
